### 2013-08-21 toc.tex

\section{Table of Contents}
\begin{enumerate}
\item Introduction (1 page) 
\item Motivation, RBF derivatives, stenciles, RBFFD  (1 page)
\item D=Ax, sparse vector matrix  multiplication
  - Ellpack format only
\item Notation (1 page)
\item Multiple derivatives of a single function
\item Single derivative of multiple functions
\item Multiple derivatives of multiple functions
\item MIC architecture (1 page)
\item Basic implementation with OpenMP (1 page)
\item MIC Vector hardware instructions, gather, masks, Pack/Unpack
  (1 page)
\item Mathematical model to predict timings, taking into account latencies, 
   cache sizes, problem size, cost of memory, cost of operations. 
\item Benchmarks. Effect of (3 pages)
\begin{itemize}
 \item Problem size
 \item Single/Double precision
 \item Bandwidth reduction
 \item Breaking up the domain into smaller pieces 
 \item Cache thrashing
 \item Indirect addressing (addressed in previous papers)
\end{itemize}
\item Domain decomposition to minimize cache thrashing
\item Bandwidth reduction, effect of domain decomposition
\item Do not address
\begin{itemize}
  \item Cost of matrix assembly
  \item Cost of accessing the derivatives in an application code (future paper)
\end{itemize}
\item Bibliography (1 page)
\end{enumerate}