### 2013-08-22 tables_and_figures.tex
Two more figures.

\section{List of Tables and Figures}
\begin{enumerate}
\item
Table: Notation for formulas
\item
Table: List of relevant vector instructions
\item
Figure: Matrix-Vector multiplication
\item
\item
\item
\item
\end{enumerate}

\section{Suggested Benchmarks}
\begin{enumerate}
\item
SpMM with 16 floats (or 8 doubles) vectors (will give upper bound in speed). Implement algorithm  from Saul \etal. 
\item
SpMM with 4 floats or 4 doubles (to compare against SpMM). As a precursor to adding more matrices. 
\item
Try to figure out why our performance is so much lower than peak. 
\end{enumerate}

\newfig{figures/rbf_stencils.pdf}{0.7}{RBFFD Stencils. Each node of the mesh is connected to $n_z-1$ stencil nodes in addition to itself. In the figure, node $A$ is connected to $B$, but $B$ is {\em not\/} connected to $A$. Thus adjacency graph of $A$ is not-symmetric. This must be corrected when reducing the bandwidth of $A$ with a Cuthill\-McKee algorithm, which assumes symmetry.}
{fig:rbf_stencils}

\newfig{figures/matrix_structure.pdf}{0.7}{Matrix Structure.}{fig:mat_struct}
\newfig{figures/swizzling.pdf}{0.7}{Duplication of the first four bytes of each channel across the entire channel.}{fig:swizzling}
\newfig{figures/channel_permutation.pdf}{0.7}{Channel Permutations on the MIC}{fig:permulatio.n}
\newfig{figures/tensor_product.pdf}{0.7}{Tensor scalar product using channels and swizzling.}{fig:tensor_product}

%\newfig{figures/theoretical_performance.png}{0.7}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors. We used a memory bandwidth of 190 Gbytes/sec (based on best measurements with specialized code) and 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:peak_perf}

\newfig{figures/gflops_peak.png}{0.7}{Peak performance of SpMV algorithm assuming either 1 or 4 matrices and/or vectors. We used a memory bandwidth of 150 Gbytes/sec (based on measurements in a code similar to the SpMV kernel, with computations removed.)}{fig:gflops_peak_perf}

\newfig{figures/speedup_wrt_base.png}{0.7}{Speedup relative to base case using one matrix and one vector.} {fig:speedkup}