### 2014-01-14 acm/aa.tex

\begin{figure}
\begin{center}
	\centering
	\begin{subfigure}[t]{0.75in}
	\centering
	\includegraphics[width=1.0\textwidth]{../figures/appendices/direct_vs_indirect_weights/compare_weight_generation/xsfc_vs_xsfc_alt_on_sph32_times_sine_20x/sph32_times_sin20x-eps-converted-to.pdf}
	\caption{Test function:  \\ $Y_{3}^{2} \sin(20 x) $.  }
		\label{fig:direct_vs_indirect_manufactured_base}
	\end{subfigure}

	\begin{subfigure}[h]{0.75in}
		\centering
	\includegraphics[width=1.0\textwidth]{../figures/appendices/direct_vs_indirect_weights/compare_weight_generation/xsfc_vs_xsfc_alt_on_sph32_times_sine_20x/pdx_sph32_times_sin20x-eps-converted-to.pdf}
	\caption{$\mathbf{p}_{x} \cdot \nabla ( Y_{3}^{2} \sin(20 x))$ }
			\label{fig:direct_vs_indirect_manufactured_xsfc}
	\end{subfigure}
	\begin{subfigure}[h]{.75in}
		\centering
	\includegraphics[width=1.0\textwidth]{../figures/appendices/direct_vs_indirect_weights/compare_weight_generation/lsfc_vs_px_grad_dot_px_grad/lsfc_sph32_times_sin20x-eps-converted-to.pdf}
	\caption{Surface Laplacian: \\ $\LaplaceBeltrami ( Y_{3}^{2} \sin(20 x) )$  }
				\label{fig:direct_vs_indirect_manufactured_lsfc}
	\end{subfigure}
	\caption{Test function and its projected derivatives on the surface of the unit sphere. }
				\label{fig:direct_vs_indirect_manufactured_solution}
	\end{center}
\end{figure}