### 2014-01-14 acm/analysis.tex

\section{Analysis}
In this section, we perform some simple analysis.

Under the best scenarios, the peak floating point performance is 2 Tflops and the peak measured bandwidth, (read: 
!80 GB/sec, write: 160 GB/sec. We seek to answer the question as to whether our algorithm is limited by the bandwidth or the floating point j
performance.

Under the best of circumstances, read occurs in $n_{rT}10^{-6}/180$ ms, writes occur in $n_{wT}10^{-6}/160$ sec. 
The calculation takes $n_F 10^{-9}/2$ ms. Many more bytes are read in than are written. Thus, we ignore the difference between the read and write 
performance.

Consider stencils of size 32, 4 vectors and 4 matrices,

The ratio of times taken by floating point operations to memory transfers is
$$
R = \frac{n_F}{n_{wT}+n_{rT}} = \frac{8}{b_x+1} \frac{10^{-6}}{160} \; \frac{2}{10^{-9}}
$$
$$
R = \frac{8}{b_x+1}  \; 10^3 80
$$

Assume $N=64$, stencils of size 32. Thus there are $250000*2*32\times 16=256 \times 10^6$ operations. At a speed of $2\times 10^{12}$ operations per sec, 
the calculation should take a minimum of $128\times 10^{-6}$ sec = 0.128 ms. Reading/writing all the data requires a transfer of 
$250000\times (32 \times 4 + 4 + 4) = 34\times 10^6$ bytes, which takes $\frac{34\times 10^6}{200\times 10^9}=0.17$ ms. 
Thus, calculation and computation take on the same order. 
Rato of calculation to transfer time is $.128/0.17=0.75 = 1/1.33$.

Now, consider the 1 vector and 1 matrix case. There are $16\times 10^6$ flops taking $8\times 10^{-6}$ sec, and a transfer of 
$250000\times (32+1+1) =8.5\times 10^6$ bytes, which takes $\frac{8.5\times 10^6}{200\times 10^9}=4.25=0.04$ ms. 
Ratio of calculation to transfer time is $.008/.04=0.2=1/5$.

Conclusion: a gain of a factor of 3.7 flops per byte, approximately 4.

If instead, one had 32 vectors, the gain would be much higher. I should implement a BENCHMARK to calculate a matrix/matrix multiplication, 
and execute it 10 times, AND make sure all elements stay in cache, to measure an idealized flop rate.

Consider A * (v,v,v,v,v,....) . Measure the performance as a function of the number of cores. The Gflop rate will measure optimum performance. 
In reality, one must take into account the cost of loop index calculations, loop indexing, etc. The hope of course, is that this cost is 
insignificant. It can be measured via loop unrolling.

TODO: need practical Gflop rate.
Use supercompact. Given 64^3 grid, as #threads increases, available cache increases, so performance shoudl go up. With low number of processors, 
I should not have enough cache (at least if using random). Using supercompact, and in addition, eliinating gather and load_epi32 so I am executing 
summation of $v_i^2$, I should get good measurement of GFlops rate. Cache will be sufficient. All I want is to avoid loads.

\subsection{measure theoretical speedup}
\subsection{measure measured speedup}
\begin{itemize}
\item register density effect: would affect flop rate
\end{itemize}