### 2014-01-17 acm/ieee_paper.tex
Incorporated Evan's changes. Blue is stuff I change (most of the stuff I cahnged). Red are questions.

%\documentclass[10pt,conference,compsocconf]{IEEEtran}
%\documentclass{acm_proc_article-sp}
\documentclass{sig-alternate}
\input{support/macros} % color is defined in macros or misc_mac
\input{support/misc_mac}
\input{support/setupicase}

\def\red#1{\textbf{\textcolor{red}{#1}}}
\def\blue#1{\textbf{\textcolor{blue}{#1}}}
\def\qes#1{{\blue{*** For Erik: #1 ***}}}
\def\es#1{{\blue{*** For Erik: #1 ***}}}
\def\ge#1{{\blue{#1}}}
\def\NOTE#1{{\red{NOTE: #1}}}
\def\ttt#1{{\tt #1}}
\def\bold#1{{\bf #1}}

\usepackage{soul}
\usepackage{xspace}
\usepackage{color} 
\definecolor{darkgreen}{rgb}{0,0.5,0}
\usepackage[colorlinks=true,% 
  linkcolor=red,% 
  citecolor=darkgreen,%
  urlcolor=blue]{hyperref}

\usepackage{listings}

\newcommand{\todo}[1]{{\color{red}\textbf{\hl{#1}}\xspace}}

\def\qes#1{}
\def\es#1{}
%\def\ge#1{}
%\usepackage{morefloats}

\newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
\newcommand{\floor}[1]{\left\lfloor#1\right\rfloor}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{graphicx}
%%\usepackage[pdftex]{graphicx}
\usepackage{subfigure}
%\usepackage{fixltx2e}
%%\usepackage{url}
%%\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\title{Acceleration of Derivative Calculations with Application to Radial Basis Function --
Finite-Differences on the Intel MIC Architecture}
%\title{Sparse Matrix Vector Multiplication with Multiple vectors and
  %Multiple Matrices on the MIC Architecture \todo{? more application oriented''}}

\subtitle{[Extended Abstract]}

%\numberofauthors{4}

%\author{
%\alignauthor
%Gordon Erlebacher\\ %\titlenote{Dr.~Trovato insisted his name be first.}\\
%       \affaddr{Department of Scientific Computing}\\
%       \affaddr{Florida State University}\\
%       \affaddr{Tallahassee, FL 32306-4120}\\
%       \email{gordon.erlebach@gmail.com}
%\alignauthor
%Erik Saule
%        \affaddr{Department of Computer Science}\\
%        \affaddr{University of North Carolina at Charlotte}\\
%        \email{esaule@uncc.edu}
%\and
%\alignauthor
%Natasha Flyer
%        \affaddr{Computational and Information System Laboratory}\\
%        \affaddr{UCAR}\\
%        \email{flyer@ucar.edu}
%\alignauthor
%Evan Bollig
%        \affaddr{Minnesota Supercomputer Institute}\\
%        \affaddr{University of Minnesota}\\
%        \affaddr{Minneapolis, MN}\\
%        \email{bollig@gmail.com}
%}
\maketitle

%\author{\IEEEauthorblockN{Gordon Erlebacher\IEEEauthorrefmark{1},
%Erik Saule\IEEEauthorrefmark{2}, Natasha Flyer\IEEEauthorrefmark{3}, 
%and Evan Bollig\IEEEauthorrefmark{1}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Scientific Computing, 
%Florida State University}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Department of Computer Science, University of North Carolina at Charlotte}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Computational and Information Systems Laboratory, UCAR}
%Email: gerlebacher@fsu.edu, esaule@uncc.edu, flyer@ucar.edu, bollig@gmail.com}
%\maketitle

\begin{abstract}
In this paper, we develop an efficient scheme for the calculation of
derivatives within the context of Radial Basis Function
Finite-Difference (RBFFD). RBF methods express functions as a linear
combination of \ge{spherically} symmetric basis functions on an arbitrary set of
nodes. The Finite-Difference component expresses this combination over
a local set of nodes neighboring the point where the derivative is
sought.  The derivative at all points takes the form of a sparse
matrix/vector multiplication (SpMV).

In this paper, we consider the case of local stencils with a fixed number
of nodes at each point and encode the sparse matrix in ELLPACK
format. We increase the number of operations relative to memory
bandwidth by \ge{interleaving} the calculation of  four derivatives 
of four different functions,
or 16 different derivatives. We demonstrate a novel implementation on
the Intel MIC architecture, taking into account its advanced swizzling and
channel interchange features. We present benchmarks on a real data
set that show an almost \ge{sevenfold} \note{eightfold; there were inconsistencies in the ppaer.} increase in speed compared to
efficient implementations of a single derivative, reaching a
performance of almost 140~Gflop/s. We explain the results through
consideration of operation count versus memory bandwidth.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{MIC, SIMD, SpMV, Sparse Matrix, Radial Basis Function}

%\begin{IEEEkeywords}
%MIC; SIMD; SpMV; SpMM; Sparse Matrix; Radial Basis Function;
%\end{IEEEkeywords}

%\IEEEpeerreviewmaketitle

\section{Introduction}
The multiplication of a sparse matrix by a dense vector (SpMV) is an
important kernel in many applied fields such as fluid dynamics 
\cite{journals/tog/BolzFGS03}, recommendation systems~\cite{Brin98} and graph
drawing~\cite{Koren05}). Naturally, improving the performance of SpMV has
captured the interest of many researchers; including the development of
various implementations for 
CPUs~\cite{Buluc2009SPAA,Williams07} and GPUs~\cite{Bell08,
  conf/ipps/KreutzerHWFBB12,
  journals/concurrency/VazquezFG11,kumar2012accelerating}. \ge{The main
challenge  to obtain good performance for matrix vector multiplication in general, and 
sparse matrix vector multiplication in particular, is the low ratio of floating point operations
to memory bandwidth. When the matrix is not dense, the problem is exacerbated due to non-uniform 
access patterns.}
%{SpMV is that
%the operations are conducted using memory locations that are irregularly spaced. 
%This makes the kernels mostly memory bound
% and there is a significant instruction overhead per flop.

% REMOVED by GE
%This makes the kernels strongly memory bound.  There is also a 
%significant instruction overhead per flop.
\NOTE{I rewrote last few sentences: write more precisely? Blas 2 is Matrix/vector, and that is still somewhat memory bound.}

Common improvement techniques such as register blocking, bandwidth
reduction (matrix reordering~\cite{Cuthill69}), partitioning to fit in
cache or TLB~\cite{Nishtala07,Temam:1992:CBS:147877.148091,conf/ppsc/Toledo97},
unrolling~\cite{Mellor-Crummey04} have impacts that are very
dependent on the matrix and overall do not lead to dramatic
improvement. (The state-of-the-art techniques in OSKI~\cite{Vuduc05}
provide some useful yet limited improvements). Register blocking~\cite{conf/ppsc/Toledo97} does
not work well to many of the matrices in general use (although it can be applied
with virtually no overhead thanks to compressed
representations~\cite{Buluc11}.) Indeed, there are about 8 bytes of the matrix to transfer
from memory 
per nonzero in single precision; each nonzero requires two
floating point operations leading to a flop-to-byte ratio of at most
$\frac{1}{4}$. This limits the obtained performance to at most a
quarter of the bandwidth of the architecture, wasting a lot of
potentially useful cycles. The commonly used techniques are mostly
designed to reach that bound rather than overcome it.

Fortunately that fate is not inevitable. One solution would be to
schedule a more instruction-intensive kernel simultaneously with the
execution of SpMV, relying on some hardware threading capabilities,
such as HyperThreading, to reduce the cycle wastage. However, most of 
the applications that use SpMV do not typically have an 
instruction-intensive kernel to run simultaneously.

Another solution, pursued in this paper, \ge{is to compute multiple SpMVs using matrices with identical spasity patterns, but with different matrix elements.}  Obviously not all the applications have such a property. However, important classes of applications such as graph recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of derivatives for solving systems of PDEs using Radial Basis Function-generated Finite Differences (RBF-FD)~\cite{FLBWSC12} \ge{fall into the category of applications that require multiple SpMVs simultaneously.} RBF-FD is a meshless method, which easily handles irregular geometries and local refinement with algorithmic complexity independent of dimensionality, 
and which can produce high-order derivative approximations. These methods are rapidly gaining 
ground in science and engineering modeling communities \cite{Bayona13,CDNT,FoL11,FLBWSC12,SPLM}). 
As a result, it is of interest to develop an efficient implementation on novel computer platforms for the calculation of derivatives within the context of RBF-FD; derivative calculations that account for the bulk of computer resources when running a numerical simulation implemented with RBF-FD. 
\ge{Typical equations solved by RBF-FD depend on multiple derivatives
of multiple functions. Rather than compute each derivative individually, this paper investigates 
how to compute all derivatives simultaneously in an interleaved fashion. Specifically, 
we calculate four different derivatives (corresponding to four different sparse matrices with identical 
sparsity pattern) of four different functions
%is an interThis paper investigates the possibility  to increase the ratio  case of approximating 
%four different derivatives, using RBF-FD, for four different functions 
(a common scenario in 3D fluid dynamics modeling) for a total of 16 derivatives. }
%Since the RBF-FD differentiation matrices approximating the four different derivative operators have an identical sparsity pattern, this leads to the simultaneous execution of 16 SpMVs at a time. Using multiple vectors at a time has been investigated before to densify the computations~\cite{Im01} but, to the best of the authors’ knowledge, this is the first time that the densification comes from adding both vectors and matrices.

To perform our analysis, we focus our attention on the improvement
that can be achieved on the Intel Xeon Phi processor. It follows the
Many Integrated Core (MIC) architecture, which has a significant
memory bandwidth and peak flop throughput thanks to its 512-bit large
SIMD registers. The Xeon Phi processor has been shown to be promising
for sparse linear algebra compared to more classical CPU or GPU
architectures~\cite{Saule13-ARXIV, Liu:2013:ESM:2464996.2465013, cramer2012openmp}.

In Section~\ref{sec:rbf}, we introduce and further motivate the RBF-FD method, giving 
an example of how calculating the derivatives for a common system of PDEs in fluid dynamics 
can be expressed as sixteen multiplications of four vectors by four sparse matrices with 
identical sparsity patterns.

Section~\ref{sec:model} presents an
estimation of the instruction intensity of various forms of the
computations. We show that a \ge{sevenfold} \NOTE{sixfold was inconsistent with eighfold in the abstract.} improvement can be
expected when computing the sixteen multiplications simultaneously to
reach a total of about 210~Gflop/s. This performance represents
approximately 10\% of the available flop/s of a Xeon Phi
coprocessor. \ge{Therefore, it is necessary to have an implementation that
performs the computation in as few instructions as
possible.} \NOTE{I DO NOT FOLLOW the Therefore in the previous sentence. Evan has the same concern.}
\ge{We describe in Section~\ref{sec:impl} the details of the MIC
architecture and how to use specialized load, store, swizzle and
permutation instructions to efficiently bring the data from memory into the vector
registers to be processed. Section~\ref{sec:expe} presents results on 
memory access and computational speed and their relation to choices made in the 
SpMV kernels.}
%experimental results  discussing the relationship between the structure of $A$ and 
%the ; specifcally on the amount of bandwidth that can be achieved
%depending on how the SpMV kernel is written. 
It also provides the actual performance of the various kernels on multiple classes of matrices, 
some generated for purpose of analysis, and some extracted from an application of RBF-FD. 
A performance of 135~Gflop/s is achieved using RBF-FD differentiation matrices. Concluding remarks and perspectives are provided in Section~\ref{sec:ccl}.\looseness=-1

\input{RBF_Section}
%\input{old_RBF_Section}

\section{Modelization of the Potential Improvements}
\label{sec:model}

We saw in the previous section that one can express the RBF problem as
a multiplication of four matrices by four vectors. We present here an
estimation of the variation on the flop intensity of the computation
and its impact on the expected performance of the
application. Relevant notation are given in Figure~\ref{tab:not}.

\begin{figure}[t]
  \begin{center}
    \scalebox{.8}{
      \begin{tabular}{|c|l|}
        %\hline
        %& & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row of $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of nonzeros\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the $n_m$ matrices in bytes\\
        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred \\
        $O$     & number of floating point operations \\
        $I_b$   & maximum computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
  \end{center}
  \caption{Notation relative to the application problem (upper section) and to the benchmarks (lower section.)}
  \label{tab:not}
\end{figure}

Each vector in the problem is of dimension $n_r$ and each entry
takes $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
  y} vectors, which lead to the size of the {\tt x} and {\tt y} vectors:
$$s_x = n_v b_x n_r\mbox{, ~}s_y = n_v n_m b_x n_r$$

The matrix is composed of $n_r$ rows and columns with $n_z$ nonzeros
per row leading to a total of $$n_c = n_r n_z$$ nonzero entries in
the matrix. Each of these nonzero entries has one index of size $b_i$
and $n_m$ values of size $b_x$. The matrices have a total size
of $$s_M = n_c (b_i + b_x n_m) = n_r n_z (b_i + b_x n_m)$$

\begin{figure*}[tbh]
  \centering
  \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
  \subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}

  \caption{Ratio of flops to bytes in single precision:  
    \ref{fig:ratio_best} best case; \ref{fig:ratio_worst} worst case when $cl=64$ 
    ; and \ref{fig:ratio_worst_nocache} worst case neglecting that the
    memory transfers are with a granularity of one cache line (equivalent to $cl = 1$)}
  \label{fig:ratio_bytes_flops}
\end{figure*}

If we assume an algorithm where the rows are processed one after
the other, the amount of memory written is precisely the
size of the {\tt y} vector. (This assumption removes the possibility of
blocking or cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$

The amount of data read from memory depends highly on both the
algorithm's execution path, and on how the matrix is structured. But
in the best case both the matrix $A$ and the source vector {\tt x} are
read once from the main memory. (We assume that all the elements of
$\tt x$ are involved in the SpMV.)  Thus, $$b_{rT} = s_M + s_x = n_r
n_z (b_i + b_x n_m) + n_v b_x n_r$$
 $$b_T = b_{rT} + b_{wT} =  n_r n_z (b_i + b_x n_m) + n_v b_x n_r (1 + n_m)$$

Notice that there is no reason for a piece of the matrix to be read
multiple times. But assuming that each element of the {\tt x} vector
is read a single time is a strong assumption. If using a single core,
it assumes that either the cache of the architecture can store the full
{\tt x} vector or that the matrix is sufficiently well structured 
to cause no cache trashing. If using multiple cores, this assumes that
no element of the {\tt x} vectors will be used by multiple
cores. \cite{Saule13-ARXIV} showed that there is very little cache trashing
in practice; however having elements of the vectors used by
multiple cores can have a significant impact on the performance
(growing with $n_v$).

On the other hand, in the worst case, every time the {\tt x} vector is
accessed, the value needs to be transferred from memory again. So in
total, there are as many transfers as the number of nonzeros in the
matrix. Note however that most architectures cannot read memory a
single byte at a time. Instead, a minimum number of bytes, equal to
the size of a cacheline $cl$, are transferred at once.  When there are
multiple vectors, each nonzero element uses $n_v$ consecutive
entries. In the worst case, the number of bytes read and transferred is
$$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$ 
$$B_T = n_v n_m b_x n_r + n_r n_z \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)$$

In SpMV, each nonzero of the matrix requires two floating point
operations: one for performing the multiplication and one for
accumulating the result row-wise. Here we are dealing with $n_v n_m$
simultaneous SpMVs and the number of floating point operations is
$$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$$

The computation intensity is the amount of floating point operations performed per
byte transferred. In the worst case and in the best case, we have
$$I_b = \frac{O}{b_T} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$$
%$$I_w = \frac{O}{B_T} = \frac{2 n_v n_m}{n_v n_m b_x n_z^{-1} + \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)}$$
$$I_w = \frac{O}{B_T} = \frac{2 n_v n_m}{(b_i+b_x n_m) + n_v n_m b_x n_z^{-1} + cl \ceil{\frac{n_vb_x}{cl}} }$$

Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios
for single precision computation in the best and the worst case on a
classical cache-based architecture ($cl = 64$) and assuming $cl=1$. We
can easily see the potential improvement in the computational
intensity when the number of matrices or vectors increases. There is a
significant difference between the best and the worst case: there is a
eightfold difference in the one matrix -- one vector ($n_m=n_v=1$) case but that gap closes
with the increase in the number of vectors and matrices to a fourfold
difference in the four matrices and four vectors ($n_m=n_v=4$) case and 1.6 fold in the 
$n_m=n_v=16$ case. 
The ratios computed with $cl = 1$ are mostly
similar to the one with $cl=64$ when the number of vectors is
large. But the differences are important when the number of vectors is
low: this highlights that accessing the memory with a granularity of one cache
line is the main problem faced when performing a standard SpMV computation.

Figure~\ref{fig:perf_predict} shows how these ratios translate
into actual performance. This figure provides projected best and worst
Gflop/s achievable assuming the computation is memory-bound and the
architecture reaches 150~GB/s. Notice that \cite{Saule13-ARXIV}
showed a higher peak bandwidth, but we will show in
Section~\ref{sec:expe} why 150~GB/s is a better estimate of what one
might achieve in these kernels. In single precision, the worse that
one could achieve by using four vectors and four matrices is much higher
than the best achievable using a classical SpMV computation. The best
performance reachable is 210~Gflop/s: almost six times higher than the
peak of the classical SpMV case.

\begin{figure}
  \centering 
  \includegraphics[width=.9\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops-peak-perf}
  \caption{Estimation of the maximum achievable performance using a
    device with a bandwidth from memory to computational units of 150~GB/s when
    varying the number of vectors and matrices.}
  \label{fig:perf_predict}
\end{figure}

\section{Efficient Implementation on the Intel Xeon Phi processor}
\label{sec:impl}

\subsection{Intel Xeon Phi}

In this work, we use an Intel Xeon Phi 5110P coprocessor. This card
has 8 memory controllers that can each execute 5 billion
transactions per second. Each has two 32-bit channels, achieving a total
bandwidth of 320~GB/s aggregated across all the memory
controllers. There are 60 cores clocked at 1.05~GHz. Their memory
interfaces are 32-bit wide with two channels and the total bandwidth
is 8.4~GB/s per core. Thus, the cores should be able to consume 504~GB/s
at most. However, the bandwidth between the cores and the memory
controllers is limited by the ring network that connects the cores and
the memory controller. Its precise bandwidth is
believed to be between 200~GB/s and 250~GB/s.

Each core in the architecture has a 32~kB L1 data cache, a 32~kB L1
instruction cache, and a 512~kB L2 cache. The architecture of a core is
based on the traditional Pentium architecture, which has been extended to
64-bit with hyperthreading. A core can hold four hardware contexts at any given time (240 threads in total). At each
clock cycle, instructions from a single thread are executed. Due to
some hardware constraints, two hardware contexts must be used to reach
the peak instruction throughput in that architecture. Similar to the
Pentium architecture, a core has two different concurrent instruction
pipelines that allow the execution of two instructions per
cycle. However, only one vector or floating point instruction can be
executed per cycle.

Most of the performance of the architecture comes from the vector
processing unit. Each core has 32 512-bit SIMD registers that can be
used for double or single precision, that is, either as a vector of eight
64-bit values or as a vector of 16 32-bit values, respectively. The
vector processing unit can perform basic instructions, such as
addition or division, and mathematical operations, such as {\tt sin} or
{\tt sqrt}, making it possible to reach 16 single precision (or 8 double precision)
operations per cycle.  
The unit also supports Fused Multiply-Add (FMA)
operations, which are typically counted as two operations 
when benchmarking. Therefore, the peak performance of the 5110P
card is one Tflop/s in double precision and two Tflop/s in single
precision. If FMA cannot be used, only half of these rates can be
achieved.

\subsection{Bringing the data into the vector register}

\begin{figure*}
  \centering
\subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{figures/code.pdf}}%
\subfigure[Content of the vector registers \ge{in the source code in (a)}.]{\hspace{-0.07\linewidth}\includegraphics[width=.5\linewidth]{figures/code-annote.pdf}}

\caption{Code snippet of the multiplication of 4 vectors with 4
  matrices. When a line changes the content of a vector register its
  new content is shown on the right. The effects of swizzling and
  permutations are highlighted using a color code.}
\label{code:mat_mul}
\end{figure*}

As explained in Section~\ref{sec:model}, we expect to achieve a performance of
about 210~Gflop/s in single precision in the best case. We will focus
on the single precision case with four vectors and four matrices. 
Similar techniques apply for other combinations. This target
performance represents 10\% of the peak performance of the
architecture, which cannot be reached without an efficient vectorization. 
Indeed, to reach such a performance, a Fused Multiply-Add
instruction on fully loaded registers must be executed at most every
10 cycles.

In term of memory layout, we adopt a format similar to the ELLPACK
format to store the matrix since the number of nonzeros per row is
constant throughout the matrix. The matrix is given in two arrays. The
first array is {\tt col\_id} that describes the
sparsity structure of the matrix. Each row of {\tt col\_id} consists
of $n_z$ entries that label the column index of each nonzero in
$A$. The other array is  {\tt data} that gives the values of
the nonzeros in the four matrices. The values of the matrices are
interleaved by groups of four so that the value $A^l_{col\_id[k]}$ of
the $k$th nonzero of the $l$th matrix is at index $\floor{\frac{k-1}{4}}*4*n_m + (l-1)*4+(k-1)\%4$.
(In other words, the first four entries are from $A^{1}$ and the 
next four are from $A^2$.) The vectors {\tt x} are
interleaved so that {\tt x}$^l_i$ and {\tt x}$^{l+1}_i$ are
consecutive in memory. The vector {\tt y} is similarly 
grouped in sets of 16, which correspond to four derivatives applied to
four functions, at a single point.\looseness=-1

We discuss how the implementation of the v4m4 kernel works using
vector instructions. The code is given for reference in
Figure~\ref{code:mat_mul}. (We will show in Section~\ref{sec:expe}
that an implementation that relies on compiler vectorization does not lead
to desirable performance, which confirms the results
of~\cite{Saule13-ARXIV}.) The registers in the MIC architecture are
512 bits wide and can store 16 floats or integers. (In comparison to
OpenCL or CUDA, one can think of these registers as a warp.) The end
goal is to load and format the nonzero values $A_h^1,A_h^2,A_h^3,A_h^4$
and the vector entries $x_m^1,x_m^2,x_m^3,x_m^4$ into two vector registers and 
apply Fused Multiply-Add on them, which is depicted in the {\tt accu} line of 
Figure~\ref{code:mat_mul}. Bringing the data into the vector
register as shown in the figure is the difficult component of our 
proposed algorithm, which we proceed to explain further. 
A thread will perform the multiplications one row at a time,
and parallelism is achieved by giving blocks of rows to each thread
using an OpenMP construct.

The core of the technique is the use of swizzling and permutation
features of the vectors in the Intel MIC architecture. A vector of 512
bits is composed of four lanes (sometimes called channels) of 128
bits, which are made of four segments of 32 bits. A swizzling
operation allows the reorder of the elements within each lane (see how
{\tt mat} is extracted from {\tt mat\_ent} in
Figure~\ref{code:mat_mul}). On the other hand, a permutation reorders entire lanes,
without changing their contents (see for instance how {\tt vect} is
generated from {\tt all\_vect} in Figure~\ref{code:mat_mul}).  Both
permutation and swizzling support common reordering and broadcast. In
the code, the ``permute'' function needs to use the {\tt
  \_mm512\_permute4f128\_epi32} instruction which is only defined on
integers and therefore in place typecasts are necessary to convert
a vector to and from integer type.  (Note that we prefer swizzling to
permutation if possible because we believe it is cheaper. Many
instructions have swizzling capability, while permutation is its own
instruction.)

Every operation that loads data from memory into a register
transfers 512 bits, or 16 floats at a time. A single nonzero
in the matrix will cause 16 operations to be performed, but it only
uses four floats from the matrix and four floats from the {\tt x} vector. So
we would like to execute a single instruction to load the data for four 
nonzeros in the matrix, and another instruction to load the data for
four nonzeros from the vectors.  
(The number of nonzeros on a row in our application is
always a multiple of four; if it weren't, we would add some explicit
zeros in the data structure for padding.)

The values from the nonzero elements in the matrix are the simplest to
load. One can load the 16 floating point values for four nonzeros of
the matrix at once. The four floats coming from each matrix are
naturally grouped within each lane of the SIMD register. One can use a
swizzle operation to broadcast the four elements of a single matrix to
fill the whole register. (The broadcast of $A_h^1,A_h^2,A_h^3,A_h^4$ is
shown in Figure~\ref{code:mat_mul}.)

Loading the entries from the {\tt x} vector is  more
involved. First, the column pointers of four different nonzeros are
loaded into a vector register and the values are distributed one per
lane of the vector register using an {\tt unpack}
operation\footnote{Notice that there are two calls to {\tt unpack} for the
LSB and MSB. Both are mandatory even if one is known to be
nilpotent according the documentation of the hardware.} (The column
pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They
are replicated so that each value fills a single vector lane 
using a swizzle operation. Then each value is multiplied by four
(because $n_v=4$) 
and the offset vector $(3,2,1,0)$ is added to each lane to obtain the
correct index of the elements of the {\tt x} vector within each lane
using a Fused Multiply-Add operation. A {\tt gather} operation is then
performed to bring the 16 entries of the {\tt x} vector into the SIMD
register. At this point we have a register where each lane is the
four vector entries for a nonzero element. Using a broadcast
permutation, one can replicate an entry of the four vectors across
each lane. (The broadcast of $x_m^1,x_m^2,x_m^3,x_m^4$ is shown in
Figure~\ref{code:mat_mul}.)

Now that $A_h^1,A_h^2,A_h^3,A_h^4$ and $x_m^1,x_m^2,x_m^3,x_m^4$ are
properly laid out in the vectors, a partial value for $y_i$ can be
computed. Once all the nonzeros of the row have been processed, the
exact $y_i$ value can be sent to memory using a the "No Read No Global
Order" hint to optimize write traffic on the memory bus.

\section{Experimental Validation}
\label{sec:expe}

\begin{figure}[t]
  \centering
  \includegraphics[width=.9\linewidth]{figures/bandwidth_read_write.pdf}

  \caption{Bandwidth performance under idealized conditions.  Entries
    with "cpp" (triangles) denote cases where vectorization is
    compiler generated.}
  \label{fig:bandwidth}
  \label{fig:band_rw_MIC}
\end{figure}

In this section, we describe a series of numerical experiments to
confirm the performance predictions of Section~\ref{sec:model}. To
better understand our results, we first perform some bandwidth
measurements, followed by the actual SpMV computations. All the codes
are compiled with the Intel C Compiler version 13 and
\mbox{-O3} optimization. The codes use the IMCI instructions exposed by
the compiler through intrinsics.

\subsection{Bandwidth}

\begin{figure}[t]
  \centering
  \includegraphics[width=.9\linewidth]{figures/bandwidth_gather_unpack.pdf}
  \caption{Performance of {\tt gather} and {\tt unpack} operations.}
  \label{fig:band_gather} 
\end{figure}

Figure~\ref{fig:bandwidth} shows a benchmark that computes the
bandwidth achieved when reading and/or writing large arrays on both
the MIC card and the CPU architecture. The read benchmark performs a
simple sum of the array. The write benchmark sets all the elemnts of
the array to zero, while the read/write benchmarks copies the array into
another one. We also investigate the difference between a straight C++
implementation (compiled with \mbox{-O3} with proper tagging to inform the
compiler of memory alignments) and a simple implementation with
explicit use of the vector registers.

The {\tt read, cpp} implementation reaches a performance of 168~GB/s
 which is better than the {\tt read}
implementation which only achieves 145~GB/s. This
is explained by the compiler adding loop unrolling and software
prefetching in the assembly code in the {\tt read, cpp} case, which is
not added in the {\tt read} case. On the other hand, the {\tt write}
case reaches a performance of 158~GB/s while the
{\tt write, cpp} implementation never exceeds 58~GB/s. 
This difference is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction, 
which bypasses the \ge{"Read For Ownership"} protocol and allows the writes
to be performed in any order. The compiler can not use this
instruction on its own because this would break the memory convention
useful for some parallel algorithms. The {\tt read/write} benchmark
performs at 140~GB/s while the {\tt read/write, cpp} benchmark performs
at about 90~GB/s.

Note that SpMV kernels are not likely to use only simple read and simple
writes. We also benchmark the performance we
can achieve using {\tt gather} and {\tt unpack} operations and we
present the results in Figure~\ref{fig:band_gather}. The {\tt unpack}
instruction can reach a performance of 140~GB/s. (There is no easy way
to implement the equivalent of the {\tt unpack} operation without
using explicit registers.) The performance of the {\tt gather}
operations depends in which order the array is read. If the indices
are ordered ({\tt compact}) then a performance of 140~GB/s is
reached. If the indices are completely randomized, the performance
drops below 5~GB/s. We can see that implementing the
indirection mechanism without vector instructions lowers the
performance significantly. (In the {\tt compact} case, it never
surpasses 100~GB/s and is close to 50~GB/s most of the time.)
\def\ww{.19\textwidth}
\begin{figure*}[t]
  \begin{center}
    \subfigure[Supercompact.]{%
      \label{fig:supercompact}
      \includegraphics[width=\ww]{figures/supercompact_matrix-crop.png}}%
    \subfigure[Compact.]{%
      \label{fig:compact}
      \includegraphics[width=\ww]{figures/compact_matrix-crop.png}}%
    \subfigure[Random.]{%
      \label{fig:random}
      \includegraphics[width=\ww]{figures/random_matrix-crop.png}}%
    %% \subfigure[2D, no RCM.]{%
    %%   \label{fig:rbf2dnorcm}
    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-norcm-crop.png}}%
    %% \subfigure[2D, RCM.]{%
    %%   \label{fig:rbf2drcm}
    %%   \includegraphics[width=\ww]{figures/kd-tree-2d-rcm-crop.png}} 
    \subfigure[3D, RCM.]{%
      \label{fig:rbf3dnorcm} 
      \includegraphics[width=\ww]{figures/kd-tree-3d-norcm-crop.png}}%
    \subfigure[3D, no RCM.]{%
      \label{fig:rbf3drcm}
      \includegraphics[width=\ww]{figures/kd-tree-3d-rcm-crop.png}}%
  \end{center}

  \caption{Different sparsity distributions. In all cases, there are
    512 rows and 32 nonzeros per row. The last two matrices
    corresponds to a derivative stencil in 3D RBFFD
    calculations (with and without RCM ordering).}
  \label{fig:spy_plots}
\end{figure*}

To summarize, using vector instructions appears necessary in order
to reach the highest bandwidth on the MIC architecture. For the instructions used
in a typical SpMV computation, one can expect a bandwidth between
140~GB/s to 160~GB/s. That is why we use in our estimations a best
achievable bandwidth of 150~GB/s despite the fact that a 
 code carefully crafted to maximize bandwidth can reach higher
values~\cite{Saule13-ARXIV}.

\subsection{Instances}

The different matrices used to perform our experiments and analysis
are presented in Figure~\ref{fig:spy_plots}. The
\ttt{Supercompact} matrices (Figure~\ref{fig:supercompact}) have been
generated to only have nonzero elements in the first 32 columns of
the matrix. As a result, only 32 values of {\tt x} are read
from memory, and the expense associated with cache misses is removed from consideration. The cost associated with 
{\tt A} remains. The \ttt{Compact} matrices
(Figure~\ref{fig:compact}) are generated to have 32 nonzeros per row
centered around the diagonal and it represents the ideal case for many
applications that rely on sparse matrix vector multiplication. The nonzero
elements of the \ttt{Random}
matrices (Figure~\ref{fig:compact}) see their nonzero elements
randomly (uniformly) distributed in the matrix; they represent the
worst case scenario for a cache-based architecture where the cache
reutilisation is the lowest from one row to the next.
The other type of matrix is used in derivative stencils in 3D
RBFFD calculations (Figure~\ref{fig:rbf3dnorcm} shows a $32$-point
stencil of a 3D $8^3$ grid).

We also apply a Reverse Cuthill-Mckee reordering to
all the matrices.  This ordering technique aims to reduce the
distance between the nonzeros and the diagonal, hopefully increasing 
the cache hit ratio.  The reordered version of the matrix can be
seen in Figure~\ref{fig:rbf3drcm}.

\vspace{-0.3em}
\subsection{Computations}

\begin{figure*}[t]
  \centering
  \subfigure[Supercompact and Random.]{\includegraphics[width=.48\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
  \subfigure[Compact and RBFFD.]{\includegraphics[width=.48\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}

  \caption{Performance of the manually vectorized code. The two
    horizontal lines depict the worst and the best predicted performances.}
  \label{fig:expe_types}
\end{figure*}

We now investigate the actual performance that we can obtain when
multiplying four vectors by four matrices in single
precision. Figure~\ref{fig:expe_types} presents results for different
types of matrices and also shows the minimum and maximum performance
predicted in Section~\ref{sec:model}. Figure~\ref{fig:perf_ideal}
gives the results for the {\tt supercompact} and {\tt random} matrices
that represent both the best and the worst case for such a practical 
computation. The {\tt supercompact} case peaks at
208~Gflop/s which is very close to the predicted peak
performance of 213~Gflop/s. Conversely the performance of the {\tt
  random} case decreases to 56~Gflop/s which is close to
the lowest predicted performance of 55~Gflop/s. One can see that RCM
ordering helps the {\tt random} case but the impact decreases when the
size of the matrix increases.

Figure~\ref{fig:perf_realistic} gives the results on the {\tt compact}
case, which represents the best realistic matrix one could find with very
structured grids and a
{\tt RBF} derivative matrix extracted from a real 3D application. In the {\tt
  compact} case, the performance can be as high as 195~Gflop/s, which is
within 15\% of the predicted  peak performance. The performance of the
{\tt RBF} case varies between 100~Gflop/s and 140~Gflop/s. Most of the
time, the RCM ordering provides an improvement which can be as high as
30~Gflop/s.

\begin{figure}[t]
  \centering

  \includegraphics[width=.9\linewidth]{figures/mic_performance_nb_threads.pdf}

  \caption{Performance of $y=Ax$ for a RBF derivative stencil of a
    $96^3$ 3D grid. Classical "one-matrix one-vector" SpMV ($n_m=n_v=1$ given in blue) has low
    performance. The manually vectorized code (red) is 3.5 times
    faster than the compiler generated one (green). RCM ordering
    (squares) only improves the performance for our manually
    vectorized code.}
  \label{fig:comparison_spmv}
  \label{fig:perf_mic}
\end{figure}

We finally investigate the central questions of this paper. Do we gain
actual performance by transforming classical SpMV computation into the
multiplication of four vectors by four matrices for the computation of the
derivative of RBFs? Does using manual vectorization improve actual
application performance? Figure~\ref{fig:comparison_spmv} compares the
performance achieved by a classical SpMV and by the multiplication of
four vectors by four matrices using either standard C++ code and our
optimized implementation. The
classical SpMV computation reaches 14~Gflop/s. 
The standard C++ implementation reaches a performance of 38~Gflop/s while our
optimized implementation almost reaches 140~Gflop/s. 
Notice that 
RCM ordering has no impact in the standard C++ implementation while it
provides a significant improvement in our manually vectorized
version. This highlights that the C++ implementation is instruction-bound
while our manually vectorized implementation is most likely 
memory-bound.

To put things in perspective, we conducted similar experiments on a
classical CPU architecture with two Intel Core i7 processors, each has eight
cores that support AVX instructions. Both the compiler-generated and
the manually-written vectorization using AVX instructions reach a
(read) bandwidth of 70~GB/s. This indicates that the performance of
the classical SpMV can not overcome 17~Gflop/s. A generic
implementation of the multiplication of four vectors with four matrices on
the {\tt RBF} matrix reaches a performance of 27~Gflop/s; while a
manual implementation using AVX instructions reaches a performance of
42~Gflop/s. This confirms that our method is sound and generalizes to
architectures other than the Intel MIC architecture.

\vspace{-0.5em}
\section{Conclusion}
\label{sec:ccl}

In the previous sections, we have explored the practical
implementation on an Intel Xeon Phi card of multiple derivative
operators acting on multiple vectors within the context of
RBFFD. Each derivative has an associated sparse matrix with a fixed
number of nonzeros on each row. While computing a single derivative of
multiple functions is rather common, we accelerate the algorithm further 
by considering multiple
matrices, with identical sparsity patterns, acting on multiple vectors. 
We specialize our study to four matrices and four derivatives,
with 16 outputs, computed as a sum of outer products.

Our implementation makes use of the IMCI MIC instruction set and
includes a number of swizzling and channel swapping operations, for an
extremely efficient tensor product implementation. \NOTE{From Evan: this is the first time $96^3$ is mentionned. Might wish to mention it earlier.} On a $96^3$ 3D
grid, it reaches a performance of 140~Gflop/s which is 2.7 times
faster than the best possible performance achievable for a single
vector and a single derivative and more than 7 times faster than a
practical implementation.

In future work, we will examine the effect of larger stencil sizes and
double precision. The presented performance does not reach the predicted
peak and we believe that common techniques from SpMV such as cache
partitioning can be successfully applied. We will also integrate the
techniques presented in this paper within a fluid simulation using
RBFFD.

\section*{Acknowledgment}
Erlebacher acknowledges funding from NSF grant
DMS-\#0934331 (FSU), and the use of the MIC at FSU and at LCSE/U. Minnesota (provided by Prof. D. Yuen). NCAR is sponsored by NSF. 
Flyer acknowledges support of NSF grant DMS-0934317.

\bibliographystyle{plain} %{IEEEabrv,paper.bib}
\bibliography{paper, bollig}
\end{document}