### 2014-06-09 slides/2014_ics/slides-ics.tex
draft of slides

\documentclass{beamer}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{epsfig}
%\usepackage{figlatex}
\usepackage{verbatim}
%\usepackage{listings}

\usepackage{subfigure}

%\usepackage{algorithm}
%\usepackage[noend]{algpseudocode}

\usepackage[ruled,vlined]{algorithm2e}

\usepackage{float}

\usetheme{Madrid}
 \setbeamertemplate{navigation symbols}{}
 \setbeamercovered{transparent}

\newcommand{\card}[1]{\ensuremath{|#1|}}
\newcommand{\union}{\ensuremath{\cup}}

\newcommand{\ceil}[1]{\left\lceil#1\right\rceil}
\newcommand{\floor}[1]{\lfloor#1\rfloor}
\newcommand{\todo}[1]{{\bf TODO: #1}}

\usealerttemplate{\color{red}\bf}{}

\graphicspath{{fig/}}

\newcommand{\MIC}{Intel MIC\xspace}

\title[RBF-FD on \MIC]{Acceleration of Derivative Calculations with
  Application to Radial Basis Function ­ Finite-Differences on the
  Intel MIC Architecture}

\date{ICS 2014}

\author[Erik Saule]{Gordon Erlebacher, {\bf Erik Saule}, Natasha Flyer, Evan
  Bollig}

\institute[UNCC]{Florida State University\\{\bf University of North Carolina at Charlotte}\\UCAR\\University of Minnesota}

\begin{document}

\maketitle

\begin{frame}
  \frametitle{Outline}
  \tableofcontents[hidesubsection,hidesubsubsection]
\end{frame}

\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\section{Introduction}

\subsection{RBF-FD}

\begin{frame}
  \frametitle{Derivative as Linear operators}
\end{frame}

\begin{frame}
  \frametitle{RBF-FD methods}

  \todo{Add definition of RBF}

  \begin{columns}
    \column{.45\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/RBFStencil_n32_a}

    An example of an RBF-FD stencil of size $n=32$ on a sphere of
    $N=1024$ nodes to approximate any derivative operator at the point
    in the square.

    \column{.55\linewidth}

    \begin{block}{Shallow waters equations of a rotating fluid}
      \tiny
      \begin{eqnarray}
        \dfrac{\partial u}{\partial t} &=& - \left( u\dfrac{\partial u}{\partial x} + v\dfrac{\partial u}{\partial y} + w\dfrac{\partial u}{\partial z} + f(yw - zv) + g\dfrac{\partial h}{\partial x} \right) 
        \nonumber \\
        \dfrac{\partial v}{\partial t} &=& - \left( u\dfrac{\partial v}{\partial x} + v\dfrac{\partial v}{\partial y} + w\dfrac{\partial v}{\partial z} + f(zu - xw) + g\dfrac{\partial h}{\partial x} \right)
        \nonumber\\
        \dfrac{\partial w}{\partial t} &=& - \left( u\dfrac{\partial w}{\partial x} + v\dfrac{\partial w}{\partial y} + w\dfrac{\partial w}{\partial z} + f(xv - yu) + g\dfrac{\partial h}{\partial x} \right)
        \nonumber\\
        \frac{\partial h}{\partial t}& =&-\left(\dfrac{\partial (uh)}{\partial x} + \dfrac{\partial (vh)}{\partial y} + h\dfrac{\partial (wh)}{\partial z}\right) , \label{height} \nonumber
      \end{eqnarray}
    \end{block}

    How to compute one simple derivative: sparse matrix
    multiplication. Bound by (roughly) 25GFlop/s. Can we do better?
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Compute all the simple derivative at once}

%  \begin{columns}
%    \column{.5\linewidth}
    \begin{block}{Idea}
      \scriptsize
      \begin{equation}
        \left( \begin{array}{cccc}
            \underline{u}_x     & \underline{v}_x     & \underline{w}_x     & \underline{h}_x \\
            \underline{u}_y     & \underline{v}_y     & \underline{w}_y     & \underline{h}_y \\
            \underline{u}_z     & \underline{v}_z     & \underline{w}_z     & \underline{h}_z \\
            \underline{u}_{hyp} & \underline{v}_{hyp} & \underline{w}_{hyp} & \underline{h}_{hyp}
          \end{array} \right)
        = \left(
          \begin{array}{c}
            D_{x}\\ D_{y}\\ D_{z}\\ D_{hyp}
          \end{array}\right)
        \times \left(\underline{u} \,\, \underline{v} \,\, \underline{w} \,\, \underline{h} \,\,\right) 
        \vspace{-1ex}
      \end{equation}
    \end{block}  
%    \column{.5\linewidth}

    \begin{block}{Why would it work}
      All the derivative matrices have the same sparsity pattern. In
      other words, the non zeros are at the same positions in the
      matrix, only the values of the non zero change.
    \end{block}

%  \end{columns}
\end{frame}

\subsection{MIC}

\begin{frame}
  \frametitle{What is Intel MIC}
  Intel Many integrated Core (MIC) is Intel's response to GPUs
  becoming popular in High Performance Computing.

  \begin{block}{What GPUs do well?}
    \begin{itemize}
      \item Get a lot of GFlops by using hundreds of cores
      \item Each core has large SIMD-like abilities
      \item Hide memory latency by using fast cycle context switch
    \end{itemize}
  \end{block}

  \begin{block}{What GPUs do not do well?}
    \begin{itemize}
      \item Alien to program
      \item Poor support for legacy applications
      \item Inter thread communications
      \item Branching
    \end{itemize}
  \end{block}

  Goal of \MIC : do all of it well!
\end{frame}

\begin{frame}
  \frametitle{Overall Architecture}
  \begin{columns}
    \column{18em}    
    \includegraphics[width=18em]{slides-figures/MIC-overview.pdf}
    \column{17em}

    \begin{itemize}
      \item 8 memory controllers
        \begin{itemize}
          \item GDDR5
          \item 2 channels (32-bit)
          \item 5.5GT/s
          \item 352GB/s aggregated peak
            \begin{itemize}
              \item twice the GPU's
              \item but you typically get half
            \end{itemize}
        \end{itemize}
      \item 50+ cores
        \begin{itemize}
          \item 32KB of L1 cache
          \item 512KB of L2 cache
          \item LRU
          \item 8-way associative
        \end{itemize}
      \item 1 PCI-e controller
        \begin{itemize}
          \item to the host (2GB/s guaranteed to memory)
        \end{itemize}
    \end{itemize}
  \end{columns}

\end{frame}

\begin{frame}
  \frametitle{Core architecture}
  \begin{columns}
    \column{20em}
    \includegraphics[width=20em,height=20em]{slides-figures/MIC-core.png}%from intelsdg  
    \column{12em}
    \begin{itemize}
      \small
    \item Clocked at 1Ghz
    \item 64-bit
    \item 4 hardware threads
      \begin{itemize}
        \footnotesize
        \item no context switching
        \item no 2 instructions for the same thread
      \end{itemize}
    \item A vectorial unit
      \begin{itemize}
        \footnotesize
        \item (more details on the next slides)
      \end{itemize}
    \item Two instruction pipes:
      \begin{itemize}
        \footnotesize
        \item 2 ALU ops
        \item ALU + MEM ops
        \item ALU + VFP ops
        \item VFP + MEM ops
      \end{itemize}
    \item In-order execution
    \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Vectorial Unit (SIMD) : The classics}

  \begin{columns}
    \column{15em}

    \includegraphics[width=15em]{slides-figures/vect-basic.png}%from intelsdg

    \column{15em}

    {\small 32 512-bits registers (16x32, 8x64)}

    \begin{block}{Binary operators (1 cycle throughput)}
      \small mul, add, sub, and, xor, or, fma
    \end{block}

    \begin{block}{Short Vector Math Library}
      sqrt, rsqrt, log, exp, div
    \end{block}

  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{When Should I Use \MIC ?}

  \begin{block}{Key points of SE10P}
    \begin{itemize}
      \item Large memory bandwidth (peak: 220GB/s)
      \item 61 cores with mandatory use of hardware threading
      \item 512-bit wide SIMD registers: 
        \begin{itemize} 
          \item FMA: up to 2x16 SP Flop/c (2x8 DP Flop/c)
          \item otherwise: up to 16 SP Flop/c (8 DP Flop/c)
        \end{itemize}
      \item On a 61-core configuration at 1.05Ghz:
        \begin{itemize} 
        \item FMA: 2x16x61x1.05Ghz = 2.048 TFlop/s SP (1.024TFlop/s DP)
        \item otherwise: 16x61x1.05Ghz = 1.024 GFlop/s SP (512GFlop/s DP)
        \end{itemize}
    \end{itemize}
  \end{block}

  Lots of bandwidth? Fused Multiply Add? Large vector registers?

  Sounds like the perfect system for Sparse Matrix operations!
\end{frame}

\begin{frame}
  \frametitle{Getting actual bandwidth}

  \begin{columns}
    \column{.49\linewidth}

    \begin{center}
      \includegraphics[width=\linewidth,page=1]{slides-figures/PPAM-data/bandwidth.pdf}

      Read
    \end{center}

    \column{.49\linewidth}

    \begin{center}
      \includegraphics[width=\linewidth,page=2]{slides-figures/PPAM-data/bandwidth.pdf}

      Write
    \end{center}

  \end{columns}

  \begin{itemize}
  \item Using the appropriate vectorial instructions gives significant
    improvements.
  \item Read Peak: 183GB/s.
  \item Write Peak: 160GB/s.
  \end{itemize}

\end{frame}

\section{Performance Prediction}

\begin{frame}
  \frametitle{Is there hope in this technique? Let's model it.}

  \begin{columns}
    \column{.4\linewidth}
    \scalebox{.55}{
      \begin{tabular}{|c|l|}
        % \hline
        % & & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row of $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of nonzeros\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the $n_m$ matrices in bytes\\
        $s_x$ & size of the $n_v$ {\tt x} vectors in bytes\\
        $s_y$ & size of the $n_v n_m$ {\tt y} vectors in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred \\
        $O$     & number of floating point operations \\
        $I_b$   & maximum computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
    \column{.6\linewidth}

    Size of the input vector:\\$s_x = n_v b_x n_r$

    Size of the matrix:\\$s_M = n_c (b_i + b_x n_m) = n_r n_z (b_i + b_x n_m)$

    Size of the write (output vector):\\$b_{wT} = s_y = n_v n_m b_x n_r$

    Data Read (best case):\\$b_{rT} = s_M + s_x$

    Data Read (worst case):\\$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}}$

    Number of floating point operation:\\$O = 2 n_v n_m n_c = 2 n_v n_m n_z n_r$
  \end{columns}

  \begin{center}
    Intensity (Best case):\\
    $I_b = \frac{O}{b_T+ b_{wT}} = \frac{2 n_v n_m}{ (b_i + b_x n_m) + n_v n_m b_x n_z^{-1} + n_v b_x n_z^{-1} }$

    Intensity (Worst case):\\
    $I_w = \frac{O}{B_T+ b_{wT}} = \frac{2 n_v n_m}{(b_i+b_x n_m) + n_v n_m b_x n_z^{-1} + cl \ceil{\frac{n_vb_x}{cl}} }$
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Ok... So will it work? Flop to byte ratio}

  \begin{columns}
    \column{.33\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_best-crop.pdf}

    \begin{center}
      Best case\\~
    \end{center}

    \column{.33\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_worst-crop.pdf}
    \begin{center}
      Worst case\\~
    \end{center}

    \column{.33\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/flops_to_bytes_no_cache-crop.pdf}
    \begin{center}
      Worst case \\no cacheline
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{In terms of flops?}

  \includegraphics[width=.9\linewidth]{slides-figures/ICS-figures/gflops_peak.pdf}

\end{frame}

\section{Implementation}

\begin{frame}
  \frametitle{Vectorial Unit (SIMD) : Rearranging}

  \begin{columns}
    \column{.45\linewidth}

    \includegraphics[width=15em]{slides-figures/vect-swizzling.png}%from intelsdg

    \column{.45\linewidth}

    Registers are organized in 4 lanes.

    \begin{block}{Permutation}
      Permutation instructions allow to reorder lanes according to
      given patterns.
    \end{block}

    \begin{block}{Swizzling}
      Almost all instructions support recombining elements within a
      lane. As it is part of the instruction, swizzling comes with no
      throughput penalty (but certainly a high latency). These also
      typically follow a given set of patterns.
    \end{block}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Vectorial Unit (SIMD) : Unaligned loads/stores}

  \begin{columns}
    \column{.45\linewidth}

    \begin{block}{Gather}
      Allows to load data from memory at arbitrary offsets. Takes one
      cycle per cache line accessed.
    \end{block}

    \includegraphics[width=\linewidth]{slides-figures/gather.pdf}

    \begin{block}{Scatter}
      Same principle as gather but does store operations.
    \end{block}

    \column{.45\linewidth}

    \begin{block}{Unpack}
      Allows to load packed data from memory using a bit mask write.
    \end{block}

    \includegraphics[width=\linewidth]{slides-figures/gather.pdf}

    \begin{block}{Pack}
      Same principle as unpack but does store operations.
    \end{block}    
  \end{columns}  
\end{frame}

\begin{frame}
  \frametitle{How to do it? (4 vectors, 4 matrices)}

  \begin{block}{Observations}
    \begin{itemize}
    \item We aim at over 200Gflop/s in single precision. (\~{}  10\% of peak.)
    \item One instruction in ten must be a Fused Multiply Add.
    \end{itemize}
  \end{block}

  \begin{figure}
    \centering
    \scalebox{.85}{
      \subfigure[Source code]{\includegraphics[clip=true,trim=4.5cm 11cm 4cm 4.2cm,width=.6\linewidth]{slides-figures/ICS-figures/code.pdf}}%
      \subfigure[Content of the vector registers.]{\hspace{-0.07\linewidth}\includegraphics[width=.55\linewidth]{slides-figures/ICS-figures/code-annote.pdf}}
    }
  \end{figure}  
\end{frame}

\section{Experiments}

\begin{frame}
  \frametitle{Results : Extreme cases}
  \begin{columns}
    \column{.4\linewidth}

    \begin{block}{Supercompact}
      \centering
      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/supercompact_matrix-crop.png}
    \end{block}

    \begin{block}{Random}
      \centering
      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/random_matrix-crop.png}
    \end{block}

    \column{.6\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/random_supercompact.pdf}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Results : Real cases}
  \begin{columns}
    \column{.4\linewidth}
    \begin{block}{Compact}
      \centering
      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/compact_matrix-crop.png}
    \end{block}

    \begin{block}{RBF-FD}
      \centering
      \includegraphics[width=.6\linewidth]{slides-figures/ICS-figures/kd-tree-3d-norcm-crop.png}
    \end{block}

    \column{.6\linewidth}
    \includegraphics[width=\linewidth]{slides-figures/ICS-figures/rbf_compact.pdf}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Does it help?}
  \begin{center}
    \includegraphics[width=.7\linewidth]{slides-figures/ICS-figures/mic_performance_nb_threads.pdf}
  \end{center}

  Performance of $y=Ax$ for a RBF derivative stencil of a $96^3$ 3D
  grid. Shown: The peaks of common methods
\end{frame}

\section{Conclusion}

\begin{frame}
  \frame{Conclusion}

  \begin{block}{Conclusion}
  \end{block}

  \begin{block}{Future Works}
    \begin{itemize}
    \item More than 1 Xeon Phi?
    \item More than 1 node?
    \item Perform a complete simulation.
    \item GPU?
    \end{itemize}
  \end{block}
\end{frame}

\bibliographystyle{alpha}
\bibliography{slides}

\end{document}